EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-04-30 05_examples.tex
Updated AHM. Should we include 2orbital with replica bath? why not...

\subsection{Bethe lattice DMFT (Fortran API)}\label{SecExamplesBetheDMFT}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
where $c^+_{i\s}$ ($c_{i\s}$) is the creation (destruction) operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and the first sum is over the nearest neighbor sites $\langle
ij\rangle$. 
The system is defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

@@

\paragraph{Results}
In the following we present some results obtained by executing this
simple program varying the interaction strenght :f:var:`uloc`.
Differently from the previous case of a quantum impurity embedded in a
given bath describing the progressive formation of a
strongly renormalized Fermi liquid state, here the DMFT
self-consistency allows to describe the transition from a correlated
metal to a Mott insulating state.

@@

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBethe.pdf}
    \caption{\label{fig1}%
        \textbf{The Mott transition.}
        }
\end{figure}

@@

The DMFT workflow for this case is very similar to the previous one
provided works in the Nambu basis defined by the spinor
$\psi_i=[c_{i\up}, c^+_{i\dw}]^t$.
In this basis any correlation functions, e.g. the Green's function, takes the matrix form:
\begin{equation}
  G =
  \begin{pmatrix}
    \hat{G}_{\uparrow\uparrow} & \hat{F}_{\uparrow\downarrow}\\
    \hat{\bar{F}}_{\downarrow\uparrow}  &    \hat{\bar{G}}_{\downarrow\downarrow} \\
   \end{pmatrix}
\end{equation}
where the symbol $\hat{A}$ indicates the possible multi-orbital
character which reduced to a scalar in the present case of a single
orbital. The symbol $\hat{\bar{A}}$ indicates that the components in
the second row are connect by particle-hole and time-reversal
symmetries to the first row one. The specific relation depends on the
symmetry of the order parameter (here $s$-wave) and the nature of the
complex argument of the functions, i.e. Matsubara or real-axis:
\begin{equation}
\begin{array}{cc}
  \hat{\bar{G}}(i\omega) = -\hat{G}^*(i\omega)\;; &  \hat{\bar{F}}(i\omega) = \hat{F}(i\omega)\\
  \hat{\bar{G}}(\omega)  = -\hat{G}^*(-\omega) \;; & \hat{\bar{F}}(\omega) = \hat{F}^*(i\omega)\\
\end{array}
\end{equation}  
The code implementation closely follows the previous structure, with
some notable exceptions related to the Nambu structure for which,
using the symmetries above, we only need to deal with the first row
independent components. In the first part of the code we initialize
the lattice structure and the solver:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipy2 import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_{loc}$ and Weiss fields

@@

The iterative scheme for the solution of DMFT closely follows the
sequence discussed in \secu{SecExamplesBetheDMFT}:   
\begin{enumerate}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed\_solve} providing the set of parameters $\vec{x}$ as input. 
\item Retrieve the self-energy functions
  $\hat{\Sigma}_{\a\b\up\up}(i\omega)$ and $\hat{S}_{\a\b\up\dw}(i\omega)$ on the 
  Matsubara axis using {\tt ed\_get\_sigma}.
\item
  Evaluate the interacting local Green's functions $\hat{G}_{loc}$ and
  $\hat{F}_{loc}$:
  \begin{equation}
  G_{loc}(i\omega) =
  \int_\RRR d\e \rho(\e)
  \begin{pmatrix}
    (i\omega +\mu)\hat{\11} - \hat{h}^0 - \hat{\Sigma}(i\omega) -\e & -\hat{S}(i\omega) \\
    -\hat{S}(i\omega) & (i\omega +\mu)\hat{\11} + \hat{h}^0 +
    \hat{\Sigma}^*(i\omega) +\e\\
  \end{pmatrix}^{-1}
\end{equation}
\item Update the normal and anomalous Weiss fields, respectively 
  $\GG^{-1}$ and $\Theta^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
    relation: $\GG^{-1}_0(i\omega) = G_{loc}(i\omega) +
    \Sigma(i\omega)$ in Nambu space. 
  \item Optimize the bath parameters $\vec{x}$ to best describe the  updated
    Weiss fields. 
\end{enumerate}
which is implemented in Python as follows:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop ):
    iloop=iloop+1
    #Solve the EFFECTIVE IMPURITY PROBLEM (first w/ a guess for the bath)
    ed.solve(bath)
    print(ed.get_phi())

    #Use aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)

    #Print...

@@

    if(iloop>1):
        bath = wmixing*bath + (1.0-wmixing)*bath_prev
    bath_prev=np.copy(bath) 
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)

@@

\paragraph{Results}
We showcase here some results for the attractive Hubbard model to
illustrate the capabilities of \NAME to solve $s$-wave superconducting
impurity problems at zero and finite temperatures. 
In figure...
\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figAHM.pdf}
    \caption{\label{fig2}%
      \textbf{The BCS to BEC crossover.}
      (a) Evolution of the spectral functions
      $-\Im{G_{loc}(\omega)}/\pi$ as a function of increasing
      attraction $U$. 
      (b) The order parameter $\phi=\langle c_\up c_\dw\rangle$ as a
      function of the attraction $U$. Data for BCS (gray) is compared
      to DMFT results (blue line and symbols). 
      (c) The correlation strength
      $\Theta=\frac{|S(0)-S(\infty)|}{S(\infty)}$ as a function of the
      attraction $U$ across the BCS-BEC crossover. 
      (d) Comparison of the double occupation $D=\langle n_\up
      n_\dw\rangle$ as a function of attraction $U$ between BCS (gray
      solid line) and DMFT (blue solid line and symbols). 
      (e) Superconducting order
      parameter $\phi$ as a function of temperature across the
      superconductor-to-normal phase transition. Data for $U=4$. The
      fit highlights the critical behavior with a mean-field exponent
      $\beta=1/2$ (red dashed line).       
        }
\end{figure}

>>>>

\subsection{Bethe lattice DMFT}\label{SecExamplesBetheDMFT}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
where $c^+_{i\s}$ ($c_{i\s}$) is the creation (destruction) operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and the first sum is over the nearest neighbor sites $\langle
ij\rangle$. 
The system is defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

@@

\paragraph{Results}
In the following we present some results obtained by executing this
simple program varying the interaction strenght $U$. 
Differently from the previous case of a quantum impurity embedded in a
given bath describing the progressive formation of a
strongly renormalized Fermi liquid state, here the DMFT
self-consistency allows to describe the transition from a correlated
metal to a Mott insulating state.

@@

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBethe.pdf}
    \caption{\label{figEx1}%
      \textbf{The Metal-Insulator Mott transition.}
      (a) Evolution of the spectral function $-\Im{G}(\omega)/\pi$ as
      a function of increasing interaction $U$. The critical
      interaction $U_c\simeq 2.8D$ separates the correlated metal $U<U_c$ from
      the Mott insulator $U>U_c$.
      (b)-(c) The corresponding evolution of the Matsubara self-energy
      $\Im\Sigma(i\omega)$ (b) and
      real-axis one $\Re\Sigma(\omega)$ across the Mott
      transition. For a particle-hole symmetric case both allow to
      estimate the renormalization constant $Z$ (see main text), using
      linear order expansion in frequency (blue solid lines). The
      values of $Z$ are reported in the legend.
      The Mott insulating solution is associated to a singularity at
      $\omega=0$ of $\Im\Sigma$. 
        }
\end{figure}

@@

The DMFT workflow for this case is very similar to the previous one
provided works in the Nambu basis defined by the spinor
$\psi_i=[c_{i\up}, c^+_{i\dw}]^t$.
In this basis any correlation functions, e.g. the Green's function, takes the matrix form:
\begin{equation}
  {\mathbf G} =
  \begin{pmatrix}
    \hat{G}_{\uparrow\uparrow} & \hat{F}_{\uparrow\downarrow}\\
    \hat{\bar{F}}_{\downarrow\uparrow}  &    \hat{\bar{G}}_{\downarrow\downarrow} \\
   \end{pmatrix}
\end{equation}
where the symbol $\hat{A}$ indicates the possible multi-orbital
character which reduced to a scalar in the present case of a single
orbital. The symbol $\hat{\bar{A}}$ indicates that the components in
the second row are connect by particle-hole and time-reversal
symmetries to the first row one. The specific relation depends on the
symmetry of the order parameter (here $s$-wave) and the nature of the
complex argument of the functions, i.e. Matsubara or real-axis:
\begin{equation}
\begin{array}{cc}
  \hat{\bar{G}}(i\omega) = -\hat{G}^*(i\omega)\;; &  \hat{\bar{F}}(i\omega) = \hat{F}(i\omega)\\
  \hat{\bar{G}}(\omega)  = -\hat{G}^*(-\omega) \;; & \hat{\bar{F}}(\omega) = \hat{F}^*(i\omega)\\
\end{array}
\end{equation}  
The code implementation closely follows the previous structure, with
some notable exceptions related to the Nambu structure for which,
using the symmetries above, we only need to deal with the first row
independent components. In the first part of the code we initialize
the lattice structure and the solver:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipy2 import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_{loc}$ and Weiss fields

@@

The iterative scheme for the solution of DMFT closely follows the
sequence discussed in \secu{SecExamplesBetheDMFT}:   
\begin{enumerate}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed.solve} providing the set of bath parameters as input. 
\item Using the dedicated
  input/output procedures {\tt ed.get\_sigma} we retrieve the self-energy functions  
  $\hat{\Sigma}(i\omega)$ and $\hat{S}(i\omega)$ on the 
  Matsubara axis.
\item
  Evaluate the interacting local Green's functions $\hat{G}_{loc}$ and
  $\hat{F}_{loc}$:
  \begin{equation}
  {\mathbf G}_{loc}(i\omega) =
  \int_\RRR d\e \rho(\e)
  \begin{pmatrix}
    (i\omega +\mu)\hat{\11} - \hat{h}^0 - \hat{\Sigma}(i\omega) -\e & -\hat{S}(i\omega) \\
    -\hat{S}(i\omega) & (i\omega +\mu)\hat{\11} + \hat{h}^0 +
    \hat{\Sigma}^*(i\omega) +\e\\
  \end{pmatrix}^{-1}
\end{equation}
\item Update the normal and anomalous Weiss fields, respectively 
  $\GG_0^{-1}$ and $\FF_0^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
    relation: $\mathbfcal{G}^{-1}_0(i\omega) = {\mathbf G}_{loc}(i\omega) +
    {\mathbf \Sigma}(i\omega)$ in Nambu space. 
  \item Optimize the bath parameters  to best describe the  updated
    Weiss fields using \NAME provided conjugate gradient  fit
    procedure {\tt ed.chi2\_fitgf}.
\end{enumerate}
which is implemented in Python as follows:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop ):
iloop=iloop+1

    #Solve impurity problem
    ed.solve(bath)

    #Self-consistency with aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)

@@

    #Error check
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)

@@

\paragraph{Results}
We showcase here some results for the DMFT solution of the attractive
Hubbard model across the BCS to BEC regime, with the aim of
illustrating the capabilities of \NAME to deal with $s$-wave
superconductivity at zero and finite temperatures.

To beging with, we report in  panel (A) of \figu{figEx2} the
evolution of the spectral density, obtained from the local normal
Green's function $-\tfrac{1}{\pi}\Im{G}_{loc}(\omega)$, as a function
of the attraction $U$. For any finite value of  $U$ the Van Hove peak
at $\omega=0$ of the 2D square lattice (see $U=0$
solution) gets split by the formation of a superconductive gap of
width which increases with $U$.
The existence of a gap is associated to the formation of a finite
order parameter $\phi=\langle c_\up c_\dw\rangle$. The behavior of
this quantity, shown in panel (B), highlights crossover from the BCS
regime at weak coupling to a BEC one for large $U$.
The former regime is characterized by the well known exponential increase of
$\phi$, which is computationally the hardest regime to adress. At
strong coupling or BEC regime the order parameter saturates to its
limiting value $\phi\to 1/2$. The comparison with mean-field result
(gray) underlines the effect of local quantum fluctuations in slightly
suppressing the order parameter. This effect is particularly active in
the intermediate regime, as captured by the behavior of the
correlation strength
$\theta=\frac{|S(i\omega\to9)-S(i\omega\to\infty)|}{S(i\omega\to\infty)}$
reported in panel (C). Assuming a monotonic behavior of the anomalous
self-energy function, this quantity estimates the ``strength'' of
quantum fluctuations comparing the mean-field-like behavior at high
energy scale with that near the Fermi level. A large value of $\theta$
indicates a more correlated system.
Our results show that both weak and strong coupling regime are
relatively weakly correlated, i.e. very near a mean-field solution. On
the contrary, the inter-mediate regime is characterized by a large
correlation, making the solution very different from that obtained in
BCS description.  
To further corroborate the distance between the DMFT and mean-field
solution we report in panel (D) the evolution of the double occupancy
$d=\langle n_\up n_\dw\rangle$. At half-filling the mean-field value is
determined solely by the order parameter as $d=1/4 + \phi^2$.
Our results show that the inclusion of quantum fluctuations in DMFT
slightly increase the value of $d$ in the weak coupling
regime. However, increasing the attraction the double occupations
quickly get reduced below the mean-field value up to the BEC regime.

Finally, we illustrate the capability of \NAME in
exploring finite temperature physics. In panel (E) we show the
temperature evolution of the order parameter $\phi(T)$ across the
superconductor-to-normal transition. The mean-field nature of this
transition is underlined by the scaling of  the solution near
the critical point $\phi \simeq (T_c-T)^\beta$ with $\beta=1/2$ as
predicted by mean-field calculations.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figAHM.pdf}
    \caption{\label{figEx2}%
      \textbf{The BCS to BEC crossover.}
      (a) Evolution of the spectral functions
      $-\Im{G_{loc}(\omega)}/\pi$ as a function of increasing
      attraction $U$. 
      (b) The order parameter $\phi=\langle c_\up c_\dw\rangle$ as a
      function of the attraction $U$. Data for BCS (gray) is compared
      to DMFT results (blue line and symbols). 
      (c) The correlation strength
      $\Theta=\frac{|S(i\omega\to0)-S(i\omega\to\infty)|}{S(i\omega\to\infty)}$ as a function of the
      attraction $U$ across the BCS-BEC crossover. 
      (d) Comparison of the double occupation $D=\langle n_\up
      n_\dw\rangle$ as a function of attraction $U$ between BCS (gray
      solid line) and DMFT (blue solid line and symbols). 
      (e) Superconducting order
      parameter $\phi$ as a function of temperature across the
      superconductor-to-normal phase transition. Data for $U=4$. The
      fit highlights the critical behavior with a mean-field exponent
      $\beta=1/2$ (red dashed line) and parameters $A\simeq 3.7$, $T_c=0.61$.       
        }
\end{figure}
